BACI of biomass change for VOCC source cells for all years

### 2019-09-10 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
apply changes to vector trimming; save plots of resulting vectors; save dataframes of matched cells to separte mulitspecies folders grouped based on vocc thresholds

Modified to allow optimal temperature to vary by maturity class. 
Therefore need to run this file separately for each species-class combination.

@@

#### Set species
```{r echo=TRUE}
# # Species run so far...
 species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

@@

#### Set vector type
```{r echo=TRUE}
 climate <- "temperature"
 temp_threshold <- c(0.5)

# climate <- "do"
# do_threshold <- c(0.5)

# climate <- "do and temperature"
```

@@

Run all subsequent code...
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
# library(gfdata)
library(sdmTMB)
# library(itsadug)
library(gfranges)
```

@@

params <- readRDS(paste0("data/optimal-values-by-species-by-age.rds"))

if (age == "imm") {
  params <- filter(params, age == "imm")
} else {
  params <- filter(params, age != "imm")
}

@@

if (climate == "temperature") {
  optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
  optimal_temps <- na.omit(optimal_temps)
  optimal_temp <- round(optimal_temps[1], digits = 2)
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  optimal_values <- optimal_temp
  max_thresholds <- temp_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do") {
  optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
  optimal_dos <- na.omit(optimal_dos)
  optimal_do <- round(optimal_dos[1], digits = 2)
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  optimal_values <- optimal_do
  min_thresholds <- do_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do and temperature") {
  optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
  optimal_temps <- na.omit(optimal_temps)
  optimal_temp <- round(optimal_temps[1], digits = 2)
  optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
  optimal_dos <- na.omit(optimal_dos)
  optimal_do <- round(optimal_dos[1], digits = 2)
  optimal_values <- c(optimal_do, optimal_temp)
  variable_names <- c("do_est", "temp")
  min_thresholds <- c(do_threshold, Inf)
  max_thresholds <- c(Inf, temp_threshold)
  indices <- c(1, 2)
  delta_t_total <- 2
  }

min_string <- paste0(min_thresholds, collapse = "-")
max_string <- paste0(max_thresholds, collapse = "-")

optimal_string <- paste0(optimal_values, collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
```

@@

  if (length(variable_names) > 1) {
    climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
  } else {
    climate_predictions <- pred_temp_do
  }
  end_time <- start_time + 2
  vocc <- make_vector_data(climate_predictions,
    variable_names = variable_names,
    ssid = model_ssid,
    start_time = start_time,
    end_time = end_time,
    skip_time = skip_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    min_thresholds = min_thresholds,
    max_thresholds = max_thresholds,
    round_fact = 10
  )

  vocc$var_1_min <- min_thresholds[1]
  vocc$var_2_min <- min_thresholds[2]
  vocc$var_1_max <- max_thresholds[1]
  vocc$var_2_max <- max_thresholds[2]

@@

vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds,
  cell_size = input_cell_size, dist_intercept = dist_intercept,
  max_dist = 120
)
})

@@

print(trimmed_vector_plots)
```
# Add species biomass data

Get all biomass predictions 
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=F}
priors <- FALSE
covariates <- "+muddy+any_rock"
covs <- gsub("\\+", "-", covariates)

@@

# Select matched control cells
```{r}
rm(vocc_by_sp)
try({
  vocc_by_sp <- readRDS(paste0(
    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})

match_formula <- paste0("treat ~ 
  muddy.st + any_rock.st + x + y") #depth.st + trawled + omega_s + 
  # for (j in (variable_names)) {
  #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  #   }

@@

  matched <- MatchIt::matchit(as.formula(match_formula), 
    exact = c("start_density.st", "start_time", "depth.st"),
    method = "nearest", m.order = "random",
    distance = "probit",
    discard="both",
    data = alldata)
  # matched
# })
}
```

Print diagnostics for matching
```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
if (!exists("vocc_by_sp")){
match_plots <- apply(matched, function(matched){  
  summary(matched)
  plot(matched)
 })
match_plots
}
```

@@

  for (j in seq_along(variable_names)) {
    climate <- paste0(variable_names[j])
    matched_time0 <- matched_time0 %>% 
      mutate((!!climate) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    matched_time1 <- matched_time1 %>% 
      mutate((!!climate) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
   }

@@

#  data
#})
vocc_by_sp <- list()
vocc_by_sp[[1]] <- data
vocc_by_sp[[3]] <- match_formula  
vocc_by_sp[[4]] <- match_plots
saveRDS(vocc_by_sp, file = paste0(
    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
}
```

```{r}
try({
  vocc_by_sp <- readRDS(paste0(
    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})
data <- vocc_by_sp [[1]]
# View(data)
```

### Plot matches
```{r}
dat_summary <- data %>% 
  group_by(start_time, cell_type) %>%
  tally() %>% filter(cell_type=="source")
data_depth <- data %>% 
  group_by(start_time) %>% 
  mutate(mean_depth = mean(depth))

@@

@@

for (j in (variable_names)) {
 gamm_formula <- paste0(gamm_formula, "+ s(std_", j, ")")
 }

>>>>

Modified to allow thresholds chosen to vary by maturity class. 
Therefore need to run this file separately for each species-class combination.

@@

#### Set species
```{r echo=TRUE}

```

@@

#### Set vector type
```{r echo=TRUE}
biomass_threshold <- "50" # filter for climate values predicted to support 50% of max biomass 
 biomass_threshold <- "25" # filter for climate values predicted to support 50% of max biomass 
global_only <- TRUE

 climate <- "temperature"
 temp_threshold <- c(0.25)

# temp_threshold <- c(0.5)
# temp_threshold <- c(0.75)
#  climate <- "do"
# #  do_threshold <- c(0.5)
#  do_threshold <- c(0.25)
# climate <- "do and temperature"
```

@@

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
# library(gfdata)
library(sdmTMB)
# library(itsadug)
library(gfranges)
```

@@

params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))

params <- params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))

if (age == "imm") {
  params <- filter(params, age == "imm")
} else {
  params <- filter(params, age != "imm")
  if (any(params$model_type == "split")) { params <- filter(params, model_type != "total") }
}

@@

if (global_only) {
 params <- params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
 params <- filter(params, global == "T")
} else {
 #params <- filter(params, ssid %in% model_ssid)
}

if (climate == "temperature") {
  # optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
  # optimal_temps <- na.omit(optimal_temps)
  # optimal_temp <- round(optimal_temps[1], digits = 2)
  # optimal_values <- optimal_temp

  if (biomass_threshold == "50") {
  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_50 }
  if (biomass_threshold == "25") {
  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_25 }
  .upper <- na.omit(.upper)

  upper_thresholds <- round(.upper[1], digits = 2)
  lower_thresholds <- NA

  variable_names <- c("temp")

  lower_change <- c(Inf)
  upper_change <- temp_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do") {
  # optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
  # optimal_dos <- na.omit(optimal_dos)
  # optimal_do <- round(optimal_dos[1], digits = 2)
  # optimal_values <- optimal_do
  params <- filter(params, lower_do_25 > 0.5)

  if (biomass_threshold == "50") {
  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_50 }
  if (biomass_threshold == "25") {
  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_25 }

  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)
  upper_thresholds <- NA
  variable_names <- c("do_est")

  lower_change <- do_threshold
  upper_change <- c(Inf)
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do and temperature") {

  # optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
  # optimal_dos <- na.omit(optimal_dos)
  # optimal_do <- round(optimal_dos[1], digits = 2)
  # optimal_values <- optimal_do

  params <- filter(params, lower_do_25 > 0.5)
  if (biomass_threshold == "50") {
  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_50 }
  if (biomass_threshold == "25") {
  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_25 }
  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)

  # optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
  # optimal_temps <- na.omit(optimal_temps)
  # optimal_temp <- round(optimal_temps[1], digits = 2)
  # optimal_values <- optimal_temp

  if (biomass_threshold == "50") {
  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_50 }
  if (biomass_threshold == "25") {
  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_25 }
  .upper <- na.omit(.upper)
  upper_thresholds <- round(.upper[1], digits = 2)

  variable_names <- c("do_est", "temp")
  lower_thresholds <- c(lower_thresholds, NA)
  upper_thresholds <- c(NA, upper_thresholds)
  lower_change <- c(do_threshold, Inf)
  upper_change <- c(Inf, temp_threshold)
  indices <- c(1, 2)
  delta_t_total <- 2
  }

trim_thresholds <- c(lower_thresholds, upper_thresholds)
min_string <- paste0(lower_change, collapse = "-")
max_string <- paste0(upper_change, collapse = "-")

optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))

glimpse(params)
```

# Run all subsequent code...

@@

  if (length(variable_names) > 1) {
    climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
  } else {
    climate_predictions <- pred_temp_do
  }
  end_time <- start_time + 2
  vocc <- make_vector_data(climate_predictions,
    variable_names = variable_names,
    ssid = model_ssid,
    start_time = start_time,
    end_time = end_time,
    skip_time = skip_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    min_thresholds = lower_change,
    max_thresholds = upper_change,
    round_fact = 10
  )

  vocc$var_1_min <- lower_change[1]
  vocc$var_2_min <- lower_change[2]
  vocc$var_1_max <- upper_change[1]
  vocc$var_2_max <- upper_change[2]

@@

vectors <- trim_vector_data(vocc, variable_names, 
  lower_change, upper_change, 
  lower_thresholds, upper_thresholds,
  cell_size = input_cell_size, dist_intercept = dist_intercept,
  max_dist = 120
)
})

@@

print(trimmed_vector_plots)
```

```{r}
png(
  file = paste0("figs/", spp,
    "/vocc-", min_string, "to", max_string, "-", 
    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = trimmed_vector_plots,
  ncol = 2,
  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
)
dev.off()
```

# Add species biomass data

Get all biomass predictions 
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=F}
priors <- FALSE
# covariates <- "+muddy+any_rock"
# covs <- gsub("\\+", "-", covariates)
covs <- gsub("\\+", "-", "-trawled+ssid")

@@

# Select matched control cells

```{r}
rm(vocc_by_sp)
try({
  vocc_by_sp <- readRDS(paste0(
    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})

#FIXME: should the cells be matched by substrate even when biomass not modelled using substrate?

match_formula <- paste0("treat ~ trawled + x + y") 
  # depth.st + omega_s + muddy.st + any_rock.st + 
  # for (j in (variable_names)) {
  #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  #   }

@@

  matched <- MatchIt::matchit(as.formula(match_formula), 
    exact = c("start_density.st", "start_time", "depth.st"),
    method = "nearest", m.order = "random",
    distance = "probit",
    discard = "both",
    data = alldata)
  # matched
# })
}
```

Print diagnostics for matching
```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
if (!exists("vocc_by_sp")){
match_plots <- plot(matched, interactive = F)
summary(matched)
}

```

@@

  for (j in seq_along(variable_names)) {
    climate_var <- paste0(variable_names[j])
    matched_time0 <- matched_time0 %>% 
      mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    matched_time1 <- matched_time1 %>% 
      mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
   }

@@

#  data
#})
vocc_by_sp <- list()
vocc_by_sp[[1]] <- data
vocc_by_sp[[3]] <- match_formula  
vocc_by_sp[[4]] <- match_plots
saveRDS(vocc_by_sp, file = paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
}
```

## Save spp dataframe
```{r}
#biomass_threshold <- "50"
optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
#climate <- "temperature"
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})

n <- nrow(vocc_by_sp[[1]])
species <- rep(species, n) 
data <- vocc_by_sp[[1]]
df <- cbind(species, data)
species <- species[[1]]

# folder to hold vocc data
biomass_threshold_string <- paste0("perc_", biomass_threshold)

dir.create(file.path("data/_all", climate_string, biomass_threshold_string))

if (climate=="temperature") dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))

if (climate=="temperature") {
  saveRDS(df, file = paste0(
    "data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", max_string,"/vocc-", climate_string, "-", 
     max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}

if (climate=="do") {
  dir.create(file.path("data/_all", climate_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string))  
  }
if (climate=="do") {
  saveRDS(df, file = paste0(
    "data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", min_string,"/vocc-", climate_string, "-",
    min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}

if (climate=="do and temperature") {
  dir.create(file.path("data/_all", climate_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
}

if (climate=="do and temperature") {
  saveRDS(df, file = paste0(
    "data/_all/", climate_string, "/", "perc_", biomass_threshold,"/vocc-", climate_string, "-",
    min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}

# View(data)
```

```{r}
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})
data <- vocc_by_sp[[1]]
```

### Plot matches
```{r}
#FIXME: SAVE PLOTS?

dat_summary <- data %>% 
  group_by(start_time, cell_type) %>%
  tally() %>% filter(cell_type=="source")
data_depth <- data %>% 
  group_by(start_time) %>% 
  mutate(mean_depth = mean(depth))

@@

```{r}
# png(
#   file = paste0("figs/", spp,
#     "/matches-", min_string, "to", max_string, "-", 
#     climate_string, "-", spp, "-", age, "-", ssid_string, "-", 
#     climate_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
#   ),
#   res = 600,
#   units = "in",
#   width = 8.5,
#   height = 11
# )
# gridExtra::grid.arrange(
#   grobs = ,
#   ncol = 2,
#   top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
# )
# dev.off()
```
```

@@

# for (j in (variable_names)) {
#  gamm_formula <- paste0(gamm_formula, "+ s(std_", j, ")")
#  }
